---
title: "Value box ideas"
format: 
  html:
    code-fold: true
    html-table-processing: none
    toc: true
    theme: custom.scss
editor: visual
execute: 
  warning: false
  error: false
  messages: false
---

```{r}
#| label: setup

library(bslib)
library(bsicons)
library(tidyverse)
library(here)
library(shiny)
library(glue)
library(knitr)
source(here("app/R/gsi_get_data.R"))
source(here("app/R/0-theme_gsi.R"))
```

This document has some options for value boxes that could appear in a row on the landing page for the dashboard.

```{r}
#| include: false
# Read in metadata
site_info <- read_csv(here("app/data/site_info.csv"))

# Download most recent data from Box
box_auth_service(token_text = Sys.getenv("BOX_TOKEN_TEXT"))
gsi_get_data()

# Read in data and join with site info
data_full <- 
  read_csv(here("app/data/gsi_living_lab_data.csv")) |> 
  right_join(site_info) |> 
  mutate(datetime = with_tz(datetime, "America/Phoenix"))
```

Filter to most recent data.

```{r}
latest_airtemp <-
  data_full |> 
  group_by(site) |> 
  select(datetime, air_temperature.value) |> 
  filter(!is.na(air_temperature.value)) |> 
  filter(datetime == max(datetime)) 

latest_precip <-
  data_full |> 
  group_by(site) |> 
  select(datetime, precipitation.value) |> 
  filter(!is.na(precipitation.value)) |> 
  filter(datetime == max(datetime)) 

latest_soiltemp <- 
  data_full |> 
  group_by(site) |> 
  select(datetime, depth_height_m, basin, soil_temperature.value) |> 
  filter(!is.na(soil_temperature.value)) |> 
  filter(datetime == max(datetime))

```

## Latest conditions in all three sites

```{r}
latest_conditions <- 
  full_join(latest_airtemp, latest_precip) |> 
  mutate(site = if_else(site == "Physics and Atmospheric Sciences", "Phys & Atm Sci", site)) |> 
  ungroup()

latest_conditions_icons <- 
  latest_conditions |> 
  select(-datetime) |> 
  mutate(
    air_temperature.value = paste("🌡️", air_temperature.value, "ºC"),
    precipitation.value = paste("🌧️", precipitation.value, "mm")
    ) 
#weird hack: if one element is <p> then the other element is large.  But if neither element is <p>, both title and value are small.  glue::glue() seems to make the title not <p>?

value_latest <- 
  value_box(
    glue::glue("Current Conditions"), 
    value = markdown(
      knitr::kable(latest_conditions_icons, col.names = c("", "", ""), format = "pipe")
      )
  )

value_latest

```

Column layout:

::: {layout-ncol="2"}
```{r}
#| echo: false
value_latest
```
```{r}
#| echo: false
value_latest
```
:::


This has good info, but in my opinion it's too much. I have to rely on a "hack" to keep the table small enough to fit in the box, and I think the font size is a bit small for a value box as a result.

## Cumulative Precip.

Cumulative precip is a little more compact

```{r}
annual_precip <-
  data_full |> 
  group_by(site) |> 
  filter(datetime >= floor_date(today(), "year")) |> 
  dplyr::summarize(
    cum_precip = sum(precipitation.value, na.rm = TRUE)
  ) |> 
  mutate(site = if_else(site == "Physics and Atmospheric Sciences", "Phys & Atm Sci", site)) |> 
  mutate(cum_precip = paste(round(cum_precip, 0), "mm"))

value_precip <- value_box(
  glue::glue("Cumulative Precipitation ({year(Sys.Date())})"),
  showcase = bs_icon("cloud-rain"),
  value = markdown(
    knitr::kable(annual_precip, col.names = c("", ""), format = "pipe")
  )
)
value_precip
```

Column layout:

::: {layout-ncol="2"}
```{r}
#| echo: false 
value_precip
```

```{r}
#| echo: false
value_precip
```
:::

## Current Air Temp.

And we can do just most recent temperature in a similar way

```{r}
airtemp <- latest_conditions |> 
  select(site, air_temperature.value) |> 
  mutate(air_temperature.value = paste(air_temperature.value, "ºC"))
latest_datetime <- format(max(latest_conditions$datetime), "%a, %b %d %I:%M %p")

value_airtemp <- value_box(
  glue::glue("{latest_datetime}"), #"hack" to remove <p> tag.  Probably a more correct way to do this
  showcase = bs_icon("thermometer-half"),
  value = markdown(
    knitr::kable(airtemp, col.names = c("", ""), format = "pipe")
  )
)
value_airtemp
```
Column layout:

::: {layout-ncol="2"}

```{r}
#| echo: false
value_airtemp
```

```{r}
#| echo: false
value_airtemp
```

:::

## Soil temperature

I'm not sure how to summarize it!  There's too many measurments

```{r}
#too many columns!
latest_soiltemp <-   
  data_full |> 
  group_by(site, depth_height_m, basin) |> 
  select(datetime, soil_temperature.value) |> 
  filter(!is.na(soil_temperature.value)) |> 
  filter(datetime == max(datetime)) |> 
  arrange(site, depth_height_m)
latest_soiltemp
```

## Single value ideas

Examples I see of value boxes generally just have a single value in them.
What about something like "coolest site" or "wettest site"?

```{r}
coolest_site <-
  latest_conditions |>
  filter(air_temperature.value == min(air_temperature.value)) 

value_coolest <-
  value_box(
    "Coolest Site",
    showcase = bs_icon("thermometer"),
    value = coolest_site$site,
    p(paste(coolest_site$air_temperature.value, "ºC")),
    p(paste("As of", format(coolest_site$datetime, "%a, %b %d %I:%M %p")))
  )
value_coolest

```

::: {layout-ncol="2"}
```{r}
#| echo: false
value_coolest
```
```{r}
#| echo: false
value_coolest
```
:::

Days since rain?

```{r}
last_rain <- data_full |> 
  select(site, datetime, precipitation.value) |> 
  filter(precipitation.value > 0) |> 
  filter(datetime == max(datetime)) |> pull(datetime)
days_since_last_rain <- (now() - last_rain) |> round() |> as.numeric()

value_last_rain <- 
  value_box(
    "Last rainfall",
    showcase = bs_icon("calendar"),
    value = p(glue::glue("{days_since_last_rain} days ago")),
    glue::glue("On {format(last_rain, '%a, %b %d')}")
  )
value_last_rain
```

Column layout:

::: {layout-ncol="2"}
```{r}
#| echo: false
value_last_rain
```
```{r}
#| echo: false
value_last_rain
```
:::

## Mix value boxes

What does it look like with two different boxes side-by-side?

::: {layout-ncol="2"}
```{r}
#| echo: false
value_airtemp
```
```{r}
#| echo: false
value_precip
```
:::


::: {layout-ncol="2"}
```{r}
#| echo: false
value_airtemp
```
```{r}
#| echo: false
value_coolest
```
:::

::: {layout-ncol="2"}
```{r}
#| echo: false
value_last_rain
```
```{r}
#| echo: false
value_coolest
```
:::


They aren't "automatically" expanded to the same size, would need to set `height` in `value_box()` I think.